---
title: 'Chapter 5: R lecture'
author: "Loren Serfass"
date: "02/24/2015"
output: html_document
---

```{r}
require(ISLR)
require(boot)
```

# Cross-validation

```{r}
?cv.glm
dim(Auto)
plot(mpg ~ horsepower, data=Auto)
glm.fit <- glm(mpg ~ horsepower, data=Auto) # This actually fits a LINEAR MODEL. (no family argument)
```

Leave-one-out CV.  Brute force makes `r dim(Auto)[1] - 1` models.

```{r}
# cv.glm(Auto, glm.fit)$delta # very slow brute force (does not use formula on page 180)
```

Using the "diagonal of the hat matrix."  Very quick!

```{r}
loocv <- function(fit) {
    hat.thing <- lm.influence(fit)$h
    mean((residuals(fit)/(1-hat.thing))^2)
}
loocv(glm.fit)
```

Trying out polynomials of different degrees. Note use of `poly` function.

```{r}
cv.error <- numeric(5)
degree <- 1:5
for (i in degree) {
    fit <- glm(mpg ~ poly(horsepower,i), data=Auto)
    cv.error[i] <- loocv(fit)
}
plot(degree, cv.error, type='b')
```

k-fold cross-validation using 10 folds.

```{r}
k <- 10
cv.error.k <- numeric(5)
for (d in degree) {
    fit <- glm(mpg ~ poly(horsepower,d), data=Auto)
    cv.error.k[d] <- cv.glm(Auto, fit, K=k)$delta[1]
}
plot(degree, cv.error, type='b')
lines(degree, cv.error.k, type='b', col='red')
```

# Bootstrap in R

Investment example from book. You have two investments, each with a variance.

Optimal formula for minimum risk:

alpha = (var(y) - cov(x,y))/(var(x) + var(y) - 2*cov(x,y))

Question: what is the sampling distribution of alpha?

```{r}
alpha <- function(x,y) {
    vx <- var(x)
    vy <- var(y)
    cxy <- cov(x,y)
    (vy - cxy) / (vx + vy - 2*cxy)
}
alpha(Portfolio$X, Portfolio$Y)
alpha.f <- function(data, index) {
    with(data[index,], alpha(X,Y))
}
alpha.f(Portfolio,1:100) # The original alpha
set.seed(1)
alpha.f(Portfolio,sample(1:100,replace=T)) # the alpha from one bootstrap resampling
boot.out <- boot(Portfolio, alpha.f, R=1000) # doing 1000 bootstrap resamplings
plot(boot.out) # two plots of the 1000 alpha statistics: histogram, and qqnorm
```

